---
title: "NavigantCode"
output: html_document
date: "2024-04-16"
---
---
title: "NavigantProject"
output: html_document
date: "2024-04-04"
---

load libraries and dataframe
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(stringr)
library(corrplot)
library(GGally)
library(reshape2)

df1 <- read_excel("DataAnalystDataSet.xls", sheet = "Demographic Data")
```
rename columns
```{r}
new_colnames <- c(
  "ZipCode",
  "Lat_deg",
  "Long_deg",
  "MedAge_24",
  "MedAge_23",
  "MedAge_22",
  "AvgAge_24",
  "AvgAge_23",
  "AvgAge_22",
  "MedHouseIncome_24",
  "MedHouseIncome_23",
  "MedHouseIncome_22",
  "AvgHouseIncome_24",
  "AvgHouseIncome_23",
  "AvgHouseIncome_22",
  "TotHouseUnits_24",
  "TotHouseUnits_23",
  "TotHouseUnits_22",
  "OwnerOccupiedHouseUnits_24",
  "OwnerOccupiedHouseUnits_23",
  "OwnerOccupiedHouseUnits_22",
  "Pop16UnempRate_24",
  "Pop16UnempRate_23",
  "Pop16UnempRate_22",
  "Pop16CLFEmp_24",
  "Pop16CLFEmp_23",
  "Pop16CLFEmp_22",
  "Pop16CLF_Unemp_24",
  "Pop16CLF_Unemp_23",
  "Pop16CLF_Unemp_22",
  "AvgCheckValue_24",
  "AvgCheckValue_23",
  "AvgCheckValue_22"
)
colnames(df1) <- new_colnames
```

```{r}
print(df1)
```

check for duplicate rows
```{r}
duplicate_rows <- duplicated(df1) | duplicated(df1, fromLast = TRUE)

print(df1[duplicate_rows, ])
```

No duplicated rows.

check column types
```{r}
column_types <- sapply(df1, class)
print(column_types)
```
Checking data type of each column. Will have to change.

```{r}
print(df1)
```
delete empty/unecessary rows
```{r}
df1 <- df1[-c(1:3), ]
print(df1)
```



load in data set
```{r}
df2 <- read_excel("DataAnalystDataSet.xls", sheet = "City Zip Data")
print(df2)
```
add in zero for zip
```{r}
df2$ZIPCD <- sprintf("%05d", df2$ZIPCD)
print(df2)
```

rename zip code column
```{r}
names(df2)[2] <- "ZipCode"
print(df2)
```
merge the datasets
```{r}
df <- merge(x=df1, y=df2, by="ZipCode")
print(df)
```

load in last data set
```{r}
df3 <- read_excel("DataAnalystDataSet.xls", sheet = "Bank Locations")
print(df3)
```

rename zip code column
```{r}
names(df3)[2] <- "ZipCode"
print(df3)
```
put zero infront of zip code
```{r}
df3$ZipCode <- sprintf("%05d", df3$ZipCode)
print(df3)
```

merge all data sets together
```{r}
aggregated_df <- df3 %>%
  group_by(ZipCode) %>%
  summarize(Banks = paste(`FI Name`, collapse = ", "),  # Include all instances
            NumBanks = n_distinct(`FI Name`))  # Count distinct banks if needed

df <- merge(df, aggregated_df, by = "ZipCode", all.x = TRUE)
```

```{r}
print(df)
```

check types of all data in complete dataframe
```{r}
str(df)
```
change necessary columns to numeric
```{r}
columns_to_convert <- c("Lat_deg", "Long_deg", "MedAge_24", "MedAge_23", "MedAge_22", "AvgAge_24", "AvgAge_23", "AvgAge_22", "MedHouseIncome_24", "MedHouseIncome_23", "MedHouseIncome_22", "AvgHouseIncome_24", "AvgHouseIncome_23", "AvgHouseIncome_22", "TotHouseUnits_24", "TotHouseUnits_23", "TotHouseUnits_22", "OwnerOccupiedHouseUnits_24", "OwnerOccupiedHouseUnits_23", "OwnerOccupiedHouseUnits_22", "Pop16CLFEmp_24", "Pop16CLFEmp_23", "Pop16CLFEmp_22", "Pop16CLF_Unemp_24", "Pop16CLF_Unemp_23", "Pop16CLF_Unemp_22", "AvgCheckValue_24", "AvgCheckValue_23", "AvgCheckValue_22", "Pop16UnempRate_24", "Pop16UnempRate_23", "Pop16UnempRate_22")

for (col in columns_to_convert) {
  df[[col]] <- as.numeric(df[[col]])
}
```

check if they changed
```{r}
str(df)
```
print names of numeric to check
```{r}
numeric_cols <- sapply(df, is.numeric)

print(names(df)[numeric_cols])
```

add in number of banks in each town column
```{r}
df$NumBanks <- ifelse(is.na(df$NumBanks), 0, df$NumBanks)
df$Banks <- ifelse(is.na(df$Banks), "none", df$Banks)
print(df)
```
checking for missing values again in full dataframe
```{r}
any(is.na(df))
```

adding in column for # of Navigant banks
```{r}
df <- df %>%
  mutate(
    NavigantCount = str_count(Banks, "Navigant")
  )
```

```{r}
head(df)
```


looping through columns for statistics, boxplots, histograms
```{r}
numeric_columns <- c("MedAge_24", "MedAge_23", "MedAge_22", "AvgAge_24", "AvgAge_23", "AvgAge_22", "MedHouseIncome_24", "MedHouseIncome_23", "MedHouseIncome_22", "AvgHouseIncome_24", "AvgHouseIncome_23", "AvgHouseIncome_22", "TotHouseUnits_24", "TotHouseUnits_23", "TotHouseUnits_22", "OwnerOccupiedHouseUnits_24", "OwnerOccupiedHouseUnits_23", "OwnerOccupiedHouseUnits_22", "Pop16CLFEmp_24", "Pop16CLFEmp_23", "Pop16CLFEmp_22", "Pop16CLF_Unemp_24", "Pop16CLF_Unemp_23", "Pop16CLF_Unemp_22", "AvgCheckValue_24", "AvgCheckValue_23", "AvgCheckValue_22", "NumBanks")

for (col in numeric_columns) {
  #summary statistics
  cat("Summary Statistics for", col, ":\n")
  print(summary(df[[col]]))
  #histogram
  hist(df[[col]], main = paste("Histogram of", col), xlab = col)
  #boxplot
  boxplot(df[[col]], main = paste("Boxplot of", col), ylab = col)
}

```
```{r}
ggplot(df, aes(x = NumBanks)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of Banks per Zip Code", x = "Number of Banks", y = "Count of Zip Codes")
```
top locations will be 10 banks per zip code and forward
bottom locations will include just zipcodes with 0 banks

Getting 7 zips with most banks and zips with 0, splitting them up to find good features from them
```{r}
top_locations <- df %>%
  arrange(desc(NumBanks)) %>%
  slice_head(n = 7)  

bottom_locations <- df %>%
  filter(NumBanks == 0)  

print(top_locations)
print(bottom_locations)

```
putting 7 zips and zips with 0 in separate vectors
```{r}
top_features <- df %>%
  filter(ZipCode %in% c("02886", "02903", "02920", "02818", "02879", "02842", "02852"))
print(top_features)

bottom_features <- df %>%
  filter(ZipCode %in% bottom_locations$ZipCode)
print(bottom_features)
```

Static Statistical Analysis

Looking at mean median range standard deviation and quantilesv and printing results
```{r}
income_columns <- c("MedAge_24","MedAge_23","MedAge_22","AvgAge_24","AvgAge_23","AvgAge_22", "MedHouseIncome_24", "MedHouseIncome_23", "MedHouseIncome_22", "AvgHouseIncome_24", "AvgHouseIncome_23", "AvgHouseIncome_22", "TotHouseUnits_24", "TotHouseUnits_23", "TotHouseUnits_22", "OwnerOccupiedHouseUnits_24", "OwnerOccupiedHouseUnits_23", "OwnerOccupiedHouseUnits_22", "Pop16CLFEmp_24", "Pop16CLFEmp_23", "Pop16CLFEmp_22", "Pop16CLF_Unemp_24", "Pop16CLF_Unemp_23", "Pop16CLF_Unemp_22", "AvgCheckValue_24", "AvgCheckValue_23", "AvgCheckValue_22")

for (column in income_columns) {
  mean_value <- mean(top_features[[column]], na.rm = TRUE)
  median_value <- median(top_features[[column]], na.rm = TRUE)
  min_value <- min(top_features[[column]], na.rm = TRUE)
  max_value <- max(top_features[[column]], na.rm = TRUE)
  range_value <- max_value - min_value #spread of data
  sd_value <- sd(top_features[[column]], na.rm = TRUE) #amount of variation
  quantiles <- quantile(top_features[[column]], probs = c(0.25, 0.75), na.rm = TRUE)
  
  cat("\nStatistics for", column, ":")
  cat("\n Mean:", mean_value)
  cat("\n Median:", median_value)
  cat("\n Min:", min_value)
  cat("\n Max:", max_value)
  cat("\n Range:", range_value)
  cat("\n Standard Deviation:", sd_value)
  cat("\n 25th Percentile:", quantiles[1])
  cat("\n 75th Percentile:", quantiles[2], "\n")
}

```
```{r}
for (column in income_columns) {
  mean_value <- mean(bottom_features[[column]], na.rm = TRUE)
  median_value <- median(bottom_features[[column]], na.rm = TRUE)
  min_value <- min(bottom_features[[column]], na.rm = TRUE)
  max_value <- max(bottom_features[[column]], na.rm = TRUE)
  range_value <- max_value - min_value #spread of data
  sd_value <- sd(bottom_features[[column]], na.rm = TRUE) #amount of variation
  quantiles <- quantile(bottom_features[[column]], probs = c(0.25, 0.75), na.rm = TRUE)
  
  cat("\nStatistics for", column, ":")
  cat("\n Mean:", mean_value)
  cat("\n Median:", median_value)
  cat("\n Min:", min_value)
  cat("\n Max:", max_value)
  cat("\n Range:", range_value)
  cat("\n Standard Deviation:", sd_value)
  cat("\n 25th Percentile:", quantiles[1])
  cat("\n 75th Percentile:", quantiles[2], "\n")
}
```
Dynamic Statistical Analysis based off key Static features
```{r}
top_features <- top_features %>%
  mutate(
    #income growth rates
    MedianIncomeGrowth_22_24 = ((MedHouseIncome_24 - MedHouseIncome_22) / MedHouseIncome_22) * 100,
    AverageIncomeGrowth_22_24 = ((AvgHouseIncome_24 - AvgHouseIncome_22) / AvgHouseIncome_22) * 100,

    #housing rates
    TotalHousingGrowth_22_24 = ((TotHouseUnits_24 - TotHouseUnits_22) / TotHouseUnits_22) * 100,
    OwnerOccupiedGrowth_22_24 = ((OwnerOccupiedHouseUnits_24 - OwnerOccupiedHouseUnits_22) / OwnerOccupiedHouseUnits_22) * 100,

    #employment growth rate
    EmploymentGrowth_22_24 = ((Pop16CLFEmp_24 - Pop16CLFEmp_22) / Pop16CLFEmp_22) * 100,

    #population growth rate
    PopulationGrowth_22_24 = (((Pop16CLFEmp_24 + Pop16CLF_Unemp_24) - (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) / (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) * 100
  )

dynamic_features <- c("MedianIncomeGrowth_22_24", "AverageIncomeGrowth_22_24", 
                      "TotalHousingGrowth_22_24", "OwnerOccupiedGrowth_22_24", 
                      "EmploymentGrowth_22_24" 
                       , "PopulationGrowth_22_24" 
                      )

for (feature in dynamic_features) {
  mean_value <- mean(top_features[[feature]], na.rm = TRUE)
  median_value <- median(top_features[[feature]], na.rm = TRUE)
  min_value <- min(top_features[[feature]], na.rm = TRUE)
  max_value <- max(top_features[[feature]], na.rm = TRUE)
  range_value <- max_value - min_value
  sd_value <- sd(top_features[[feature]], na.rm = TRUE)
  quantiles <- quantile(top_features[[feature]], probs = c(0.25, 0.75), na.rm = TRUE)

  cat("\nStatistics for", feature, ":")
  cat("\n Mean:", mean_value)
  cat("\n Median:", median_value)
  cat("\n Min:", min_value)
  cat("\n Max:", max_value)
  cat("\n Range:", range_value)
  cat("\n Standard Deviation:", sd_value)
  cat("\n 25th Percentile:", quantiles[1])
  cat("\n 75th Percentile:", quantiles[2], "\n")
}

```



```{r}
bottom_features <- bottom_features %>%
  mutate(
    #income growth rates
    MedianIncomeGrowth_22_24 = ((MedHouseIncome_24 - MedHouseIncome_22) / MedHouseIncome_22) * 100,
    AverageIncomeGrowth_22_24 = ((AvgHouseIncome_24 - AvgHouseIncome_22) / AvgHouseIncome_22) * 100,

    #housing rates
    TotalHousingGrowth_22_24 = ((TotHouseUnits_24 - TotHouseUnits_22) / TotHouseUnits_22) * 100,
    OwnerOccupiedGrowth_22_24 = ((OwnerOccupiedHouseUnits_24 - OwnerOccupiedHouseUnits_22) / OwnerOccupiedHouseUnits_22) * 100,

    #employment growth rate
    EmploymentGrowth_22_24 = ((Pop16CLFEmp_24 - Pop16CLFEmp_22) / Pop16CLFEmp_22) * 100,

    #population growth rate
    PopulationGrowth_22_24 = (((Pop16CLFEmp_24 + Pop16CLF_Unemp_24) - (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) / (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) * 100
  )

dynamic_features <- c("MedianIncomeGrowth_22_24", "AverageIncomeGrowth_22_24", 
                      "TotalHousingGrowth_22_24", "OwnerOccupiedGrowth_22_24", 
                      "EmploymentGrowth_22_24" 
                       , "PopulationGrowth_22_24" 
                      )

for (feature in dynamic_features) {
  mean_value <- mean(bottom_features[[feature]], na.rm = TRUE)
  median_value <- median(bottom_features[[feature]], na.rm = TRUE)
  min_value <- min(bottom_features[[feature]], na.rm = TRUE)
  max_value <- max(bottom_features[[feature]], na.rm = TRUE)
  range_value <- max_value - min_value
  sd_value <- sd(bottom_features[[feature]], na.rm = TRUE)
  quantiles <- quantile(bottom_features[[feature]], probs = c(0.25, 0.75), na.rm = TRUE)

  cat("\nStatistics for", feature, ":")
  cat("\n Mean:", mean_value)
  cat("\n Median:", median_value)
  cat("\n Min:", min_value)
  cat("\n Max:", max_value)
  cat("\n Range:", range_value)
  cat("\n Standard Deviation:", sd_value)
  cat("\n 25th Percentile:", quantiles[1])
  cat("\n 75th Percentile:", quantiles[2], "\n")
}

```
```{r}
head(df)
```

```{r}
df <- df %>%
  mutate(
    #static features
    MedianHouseIncome_2024 = MedHouseIncome_24,  
    AverageHouseIncome_2024 = AvgHouseIncome_24,  
    TotalHousingUnits_2024 = TotHouseUnits_24,  
    OwnerOccupiedUnits_2024 = OwnerOccupiedHouseUnits_24,  
    EmploymentRate_2024 = Pop16CLFEmp_24 / (Pop16CLFEmp_24 + Pop16CLF_Unemp_24),  
    Population_2024 = Pop16CLFEmp_24 + Pop16CLF_Unemp_24,  
    MedianAge_2024 = MedAge_24,  

    # dynamic features - rates from 2022 to 2024
    MedianIncomeGrowthRate = (MedHouseIncome_24 - MedHouseIncome_22) / MedHouseIncome_22 * 100,
    AverageIncomeGrowthRate = (AvgHouseIncome_24 - AvgHouseIncome_22) / AvgHouseIncome_22 * 100,
    HousingGrowthRate = (TotHouseUnits_24 - TotHouseUnits_22) / TotHouseUnits_22 * 100,
    OwnerOccupiedGrowthRate = (OwnerOccupiedHouseUnits_24 - OwnerOccupiedHouseUnits_22) / OwnerOccupiedHouseUnits_22 * 100,
    EmploymentGrowthRate = (Pop16CLFEmp_24 - Pop16CLFEmp_22) / Pop16CLFEmp_22 * 100,
    PopulationGrowthRate = (((Pop16CLFEmp_24 + Pop16CLF_Unemp_24) - (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) / (Pop16CLFEmp_22 + Pop16CLF_Unemp_22)) * 100 
  )

print(df)

```
creating new data frame using features created
```{r}
analysis_df <- df[, c("ZipCode", "MedianHouseIncome_2024", "AverageHouseIncome_2024", "TotalHousingUnits_2024", "OwnerOccupiedUnits_2024", "EmploymentRate_2024", "Population_2024", "MedianAge_2024", "MedianIncomeGrowthRate", "AverageIncomeGrowthRate", "HousingGrowthRate", "OwnerOccupiedGrowthRate", "EmploymentGrowthRate", "PopulationGrowthRate")]
head(analysis_df)
```

statistics of new features
```{r}
print_column_stats <- function(data, column_name) {
  if (!is.numeric(data[[column_name]])) {
    cat("\n---", column_name, "---\n")
    cat("non-numeric column\n")
  } else {
    column_data <- data[[column_name]]
    cat("\n---", column_name, "---\n")
    cat("Mean:", mean(column_data, na.rm = TRUE), "\n")
    cat("Median:", median(column_data, na.rm = TRUE), "\n")
    cat("Standard Deviation:", sd(column_data, na.rm = TRUE), "\n")
    cat("Min:", min(column_data, na.rm = TRUE), "\n")
    cat("Max:", max(column_data, na.rm = TRUE), "\n")
    quantiles <- quantile(column_data, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
    cat("25th Percentile:", quantiles[1], "\n")
    cat("75th Percentile:", quantiles[3], "\n")
    cat("Summary:\n")
    print(summary(column_data))
  }
}

sapply(names(analysis_df), function(col) print_column_stats(analysis_df, col))
```
PCA
```{r}
#scaling
analysis_df_scaled <- analysis_df %>%
  select(MedianHouseIncome_2024, AverageHouseIncome_2024, TotalHousingUnits_2024, 
         OwnerOccupiedUnits_2024, EmploymentRate_2024, Population_2024, 
         MedianAge_2024, MedianIncomeGrowthRate, AverageIncomeGrowthRate, 
         HousingGrowthRate, OwnerOccupiedGrowthRate, EmploymentGrowthRate, 
         PopulationGrowthRate) %>%
  scale()

#PCA
pca_result <- prcomp(analysis_df_scaled, center = TRUE, scale. = TRUE)

# visualization
library(factoextra)
fviz_pca_biplot(pca_result, label = "var", geom.ind = "point", col.ind = "black", 
                col.var = "#2E9FDF", addEllipses = TRUE)


```

finding top/bottom groups by using scoring model on pca loadings
```{r}
loadings <- pca_result$rotation[, 1]

analysis_df$Score <- as.matrix(analysis_df_scaled) %*% loadings

sorted_zip_codes <- analysis_df %>%
  arrange(desc(Score)) %>%
  mutate(Rank = row_number())

top_zip_codes <- head(sorted_zip_codes, 15)  # Top 15
bottom_zip_codes <- tail(sorted_zip_codes, 15)  # Bottom 15
```

```{r}
common_zip_codes <- intersect(top_zip_codes$ZipCode, bottom_zip_codes$ZipCode)
if (length(common_zip_codes) > 0) {
}

analysis_df$Group <- with(analysis_df, ifelse(ZipCode %in% top_zip_codes$ZipCode, "Top Group", ifelse(ZipCode %in% bottom_zip_codes$ZipCode, "Bottom Group", "Other")))

```


ANOVA and Tukey
```{r}
features <- c("MedianHouseIncome_2024", "AverageHouseIncome_2024", "TotalHousingUnits_2024", 
              "OwnerOccupiedUnits_2024", "EmploymentRate_2024", "Population_2024", 
              "MedianAge_2024", "MedianIncomeGrowthRate", "AverageIncomeGrowthRate", 
              "HousingGrowthRate", "OwnerOccupiedGrowthRate", "EmploymentGrowthRate", 
              "PopulationGrowthRate")

run_anova_and_tukey <- function(data, feature) {
  anova_result <- aov(reformulate("Group", response = feature), data = data)
  anova_summary <- summary(anova_result)
  
  cat("\nANOVA results for", feature, ":\n")
  print(anova_summary)
  
  if (anova_summary[[1]]$"Pr(>F)"[1] < 0.05) {
    tukey_result <- TukeyHSD(anova_result)
    cat("\nTukey HSD results for significant ANOVA of", feature, ":\n")
    print(tukey_result)
  } else {
    cat("\nNo significant differences found for", feature, "\n")
  }
}

lapply(features, function(f) run_anova_and_tukey(analysis_df, f))

```

visualizing PCA
```{r}
ggplot(analysis_df, aes(x = Score, fill = Group)) +
  geom_histogram(bins = 30, position = "identity", alpha = 0.5) +
  scale_fill_manual(values = c("Top Group" = "green", "Bottom Group" = "red", "Other" = "grey")) +
  labs(title = "Distribution of PCA Scores by Group", x = "PCA Score", y = "Count")

```

Heat Map to see correlated variables
```{r}
cor_matrix <- cor(significant_data, use = "complete.obs")

melted_cor_matrix <- melt(cor_matrix)

ggplot(melted_cor_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
        axis.text.y = element_text(size = 12)) +
  labs(x = "", y = "", title = "Correlation Matrix Heatmap") +
  coord_fixed()



```

first one based off correlation (variables selected based off heat map, weights chosen from what I think is most important)


```{r}
selected_weights <- c(
  MedianHouseIncome_2024 = 0.30,
  OwnerOccupiedUnits_2024 = 0.20,
  EmploymentGrowthRate = 0.15,
  Population_2024 = 0.20,  # Only if distinctly necessary
  MedianIncomeGrowthRate = 0.15
)

analysis_df$Score <- with(analysis_df, (
  MedianHouseIncome_2024 * selected_weights["MedianHouseIncome_2024"] +
  OwnerOccupiedUnits_2024 * selected_weights["OwnerOccupiedUnits_2024"] +
  EmploymentGrowthRate * selected_weights["EmploymentGrowthRate"] +
  Population_2024 * selected_weights["Population_2024"] +  # If retained
  MedianIncomeGrowthRate * selected_weights["MedianIncomeGrowthRate"]
))

analysis_df <- analysis_df %>%
  arrange(desc(Score)) %>%
  mutate(rank = row_number())

top_zip_codes <- head(analysis_df, 10)
bottom_zip_codes <- tail(analysis_df, 10)

print("Top Zip Codes Based on Score:")
print(top_zip_codes)

print("Bottom Zip Codes Based on Score:")
print(bottom_zip_codes)


```

second one new score
```{r}
new_weights <- c(
  MedianHouseIncome_2024 = 0.3,
  TotalHousingUnits_2024 = 0.25,
  OwnerOccupiedUnits_2024 = 0.25,
  Population_2024 = 0.2
)

analysis_df$New_Score <- with(analysis_df, (
  MedianHouseIncome_2024 * new_weights["MedianHouseIncome_2024"] +
  TotalHousingUnits_2024 * new_weights["TotalHousingUnits_2024"] +
  OwnerOccupiedUnits_2024 * new_weights["OwnerOccupiedUnits_2024"] +
  Population_2024 * new_weights["Population_2024"]
))

analysis_df <- analysis_df %>%
  arrange(desc(New_Score)) %>%
  mutate(New_Rank = row_number())

```
 
```{r}
top_zip_codes_new <- head(analysis_df, 10)
bottom_zip_codes_new <- tail(analysis_df, 10)

print("Top Zip Codes Based on New Scores:")
print(top_zip_codes_new)

print("Bottom Zip Codes Based on New Scores:")
print(bottom_zip_codes_new)

```

third one
```{r}
selected_weights <- c(
  MedianHouseIncome_2024 = 0.20,
  AverageHouseIncome_2024 = 0.20,
  Population_2024 = 0.20,
  PopulationGrowthRate = 0.15,
  EmploymentGrowthRate = 0.15,
  OwnerOccupiedUnits_2024 = 0.10
)

analysis_df$Score <- with(analysis_df, (
  MedianHouseIncome_2024 * selected_weights["MedianHouseIncome_2024"] +
  AverageHouseIncome_2024 * selected_weights["AverageHouseIncome_2024"] +
  Population_2024 * selected_weights["Population_2024"] +
  PopulationGrowthRate * selected_weights["PopulationGrowthRate"] +
  EmploymentGrowthRate * selected_weights["EmploymentGrowthRate"] +
  OwnerOccupiedUnits_2024 * selected_weights["OwnerOccupiedUnits_2024"]
))

analysis_df <- analysis_df %>%
  arrange(desc(Score)) %>%
  mutate(Rank = row_number())

top_zip_codes <- head(analysis_df, 15)
bottom_zip_codes <- tail(analysis_df, 15)

print("Top Zip Codes Based on Score:")
print(top_zip_codes)

print("Bottom Zip Codes Based on Score:")
print(bottom_zip_codes)

```

confused because some of these for the top/bottom locations when looking at the group they were put in previously, the group will have a lot of top locations in the bottom zip codes and vice versa