---
title: "Assignment 7: K-means Clustering"
format: 
  html: 
    toc: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

------------------------------------------------------------------------

Given the following dataset

```{r}
library(tidyverse)
x = c(1,2,3,4,5)
set.seed(1983)
y = sample(1:10, replace = TRUE, 5)
  
df <- data.frame(Cluster = c(1,2,1,1,2),
                 Point = LETTERS[1:5],
                 x = x,
                 y = y)
df = as_tibble(df)

library(knitr)
kable(df %>% select(-Cluster))
```
1. Consider all the points are in one cluster.  Find the centroid of the data. Calculate the total variance within of the cluster. 

2. Let assign A and B for cluster 1 and the remainders for cluster 2. Find the two centroids of the two clusters and calculate the total variance within of the cluster. 

3. Calculate the distance of all the points to the two centroids. Reassign the points to the appropriate cluster.
